---
title: "Assignment2"
author: "John Shen","Elaina Sinclair","Zihao Liu"
date: "2/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

This statement is problematic because if the correlation does not reflect true cause, it would be a spurious correlation - and treating it like a causal effect may not lead to improvement in health in the future. 

## Question 2

First number = E (exposure), second number = A (adherence)
risk difference: $$E(Y^{11} - Y^{01})$$
risk ratio: $$\frac{E(Y^{11})}{E(Y^{01})}$$
Odds ratio: $$\frac{OddsE(Y^{11})}{OddsE(Y^{01})}$$

## Question 3

Difference between the risk if the entire treated population would receive four consecutive days (days 1-4) of treatment versus the risk if the entire treated population would received only three consecutive days of treatment (days 1-3).

## Question 4

$$\frac{E(Y^{d_{4}=1})}{E(Y^{d_{4}=0})}$$

## Question 5



## Question 6

Risk difference: If all smokers quit smoking they would have a 12.9% reduction in risk of high blood pressure (12.9 out of 100 fewer former smokers would not develop high blood pressure)
Risk  Ratio: If all smokers quit smoking, their risk of high blood pressure would be 53% of the risk if they had continued to smoke.
Odds Ratio: If all smokers quit smoking, their odds of high blood pressure would be 45% of the odds if the had continued to smoke.

## Question 7

The average treatment effect would be more relevant for the relation between quitting smoking and high blood pressure because we are interested in how quitting smoking would affect all smokers. If we only look at effect of quitting smoking among those who were observed to have quit smoking (i.e. effect of treatment on the treated), there may be other confounding factors (e.g. those who quit may tend to engage in more healthier lifestyle changes than those who didn't) that would affect their blood pressure and alter effect of quitting smoking on this subgroup. 

## Question 8

Participant who would still go through second-hand smoking even after he or she quit smoking already, thus the potensial outcome would be affected y the exposure status of other individual.

## Question 9

The authors are assuming consistency and no interference (SUTVA) because these two assumptions are what allow us to equate the observed outcomes conditioned on the treatment to the potential outcomes conditioned on the treatment.

## Question 10

The word "exchangeable" is used because you are able to exchange the exposed and unexposed groups without changing the effect of the exposure, in other words, the effect of exposure on the outcome does not depend on which group receives the exposure. 

## Question 11

The total number of possible interactions in the model is 4083 because according to the formula provided in the notes, with 12 variables there are 2^12 - 12 - 1 possible interactions. We have verified this formula by summing the number of combinations for 12 choose 2, 12 choose 3, ..., 12 choose 12 (12 choose 0 and 12 choose 1 are excluded as interactions must be between at least 2 variables). The total number of 2-way interactions is the total number of combination of 2 variables which can be calculated using the equation 12!/(10!*2!) = 66 2-way interactions.

## Question 12

Yes, $$\theta$$ is 0 if exchangeability holds because there is no association between outcome and exposure and thus we would expect that the result of the odds ratio be a null result, 1 (exponentiating 0 is equal to 1). However, if exchangeability does not hold, we know $$\theta$$ is not 0 but it could be any other number. 


**Question 13:** Install the AIPW package from CRAN and load the library. Then load the "eager_sim_obs" dataset using the data("eager_sim_obs") command. Using these data, generate (i) a propensity score over lap plot, (ii) a list of the five largest and smallest propensity scores, and (iii) the summary distribution of the stabilized inverse probability weights using a propensity score model that adjusts for eligibility, loss_num, age, time_try_pregnant, BMI, and meanAP. Is positivity violated in these data? Why or why not?

ANS: The mean of the stabilized weights is 1, and the max weight is not large at all, suggesting very well-behaved weights. Thus, in this particular case, we are not concerned with violations of the positivity assumption.

```{r}
library(AIPW)
library(dplyr)
library(ggplot2)
data("eager_sim_obs")

propensity_score <- glm(sim_A ~ eligibility + loss_num + age + time_try_pregnant+BMI+meanAP, data = eager_sim_obs, family = binomial(link = "logit"))$fitted.values


set.seed(123)
exposure <- eager_sim_obs$sim_A
plot_data <- data.frame(propensity_score,Exposure = as.factor(eager_sim_obs$sim_A))
                        
ggplot(data = plot_data) + 
  scale_y_continuous(expand = c(0,0)) + 
  scale_x_continuous(expand = c(0,0)) + 
  ylab("Density") + 
  xlab("Propensity Score") +
  scale_color_manual(values = c("#000000","#D55E00")) + 
  geom_density(aes(x = propensity_score, group = Exposure, color = Exposure)) +
  xlim(0, 1)


pscore<-as.data.frame(propensity_score)

high_five<-pscore %>% 
  slice_max(order_by = propensity_score, n = 5) %>% 
  rename(High=propensity_score)

low_five<-pscore %>% 
  slice_min(order_by = propensity_score, n = 5) %>% 
  rename(Low=propensity_score)

cbind(high_five,low_five)


sw <- (mean(exposure)/propensity_score) *exposure + ((1 - mean(exposure))/(1-propensity_score)) * (1 - exposure)
summary(sw)
```


**Question 14:** Consider a two-arm placebo controlled randomized trial with four mutually exclusive strata labeled S = 1,S = 2,S = 3 and S = 4. Suppose that the treatment was assigned to: 20% of individuals in stratum S = 1; 30% of individuals in stratum S = 2; 15% of individuals in stratum S = 3; and 10% of individuals in stratum S = 4. Can you determine all of the propensity score values in the sample of individuals in the trial?

ANS: Yes



**(Bonus?) Question 15** Using the information provided in Question 14, please write a logistic regression equation that the defines the propensity score for this randomized trial. What are the parameter values in this logistic regression model?

```{r}
propensity_score <- glm(Treatment~s1+s2+s3 , data = df1, family = binomial(link = "logit"))$fitted.values
exposure <- B0(intercept) + B1(dummy var) + B2(dummy var) + B3(dummy var)
```


